A preliminary constraint grammar for Russian

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
revising more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60220 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{times}
\usepackage{latexsym}
\usepackage{fixltx2e} %allows subscripts
\usepackage{tipa} %allows IPA symbols
\usepackage{linguex}
\usepackage[multiple]{footmisc} % separates consecutive footnotes with comma
%\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage{alltt}
\usepackage{graphicx}
\usepackage[small,bf]{caption}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

@@

We identify three different types of morphosyntactic ambiguity: 
intraparadigmatic, morphosyntactically incongruent, and morphosyntactically 
congruent. The following examples make use of stress ambiguity to illustrate 
each kind of ambiguity.\footnote{Written standard Russian does not typically
indicate stress position.}\footnote{A recent study by \newcite{reynolds.tyers-15} 
found that about 7.5\% of 
morphosyntactic ambiguity in a corpus of Russian resulted in stress placement 
ambiguity.} \emph{Intraparadigmatic} ambiguity refers to homographic 
wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.

@@

\emph{Morphosyntactically congruent} ambiguity occurs between homographs 
that belong to separate lexemes, and whose morphosyntactic values are identical, 
as shown in \ref{ex:MScongruent}. As indicated in Table~\ref{table:ambiguity}, 
this kind of ambiguity is relatively rare, and resolving this ambiguity requires 
the use of technologies such as word sense disambiguation.

@@

Table~\ref{table:ambiguity} shows that most morphosyntactic ambiguity in 
unrestricted Russian text is rooted in intraparadigmatic and morphosyntactically 
incongruent ambiguity. Detailed part-of-speech tagging with morphosyntactic 
analysis can help disambiguate these forms. However, morphosyntactically congruent
ambiguity cannot be disambiguated by means of detailed part-of-speech tagging, 
and since congruent stress ambiguity represents only a small percentage of 
ambiguous wordforms, we leave it to future work.

@@

\ex. \label{ex:FSToutput} 
	\a. \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}
	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}
%	\c. 
% det/prn (e.g. pos... их, я, мой)
% 'over'generation (e.g. all dets get all gen/nbr/case combinations)
%^^^
%ROB

@@

The constraint grammar is composed of 299 rules which are divided into four
categories: Safe, Safe heuristic, Heuristic, and Syntax labelling. The distribution
of rules is shown in Table~\ref{tab:ruleDist}.

@@

The philosophy is that safe rules should represent real constraints in
the language. Examples might be that a preposition cannot directly precede a finite verb
or that prepositional case requires a preceeding preposition.

>>>>

\usepackage{times}
\usepackage{latexsym}
\usepackage{fixltx2e} %allows subscripts
\usepackage{tipa} %allows IPA symbols
\usepackage{linguex}
%\usepackage[multiple]{footmisc} % separates consecutive footnotes with comma
%\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage{alltt}
\usepackage{graphicx}
\usepackage[small,bf]{caption}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

@@

We identify three different types of morphosyntactic ambiguity: 
intraparadigmatic, morphosyntactically incongruent, and morphosyntactically 
congruent. The following examples make use of word stress ambiguity to illustrate 
each kind of ambiguity.\footnote{Written standard Russian does not typically
indicate stress position, but knowing stress position is essential for pronunciation. 
A recent study by \newcite{reynolds.tyers-15} found that about 7.5\% of 
morphosyntactic ambiguity in a corpus of Russian resulted in stress position 
ambiguity.} \emph{Intraparadigmatic} ambiguity refers to homographic 
wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.

@@

\emph{Morphosyntactically congruent} ambiguity occurs between homographs 
that belong to separate lexemes, and whose morphosyntactic values are identical, 
as shown in \ref{ex:MScongruent}.

@@

Table~\ref{table:ambiguity} shows that most morphosyntactic ambiguity in 
unrestricted Russian text is rooted in intraparadigmatic and morphosyntactically 
incongruent ambiguity. Detailed part-of-speech tagging with morphosyntactic 
analysis can help disambiguate these forms. On the other hand, morphosyntactically congruent
ambiguity represents only a very small percentage of ambiguous wordforms, and instead 
of detailed part-of-speech tagging, it can be resolved by means of word sense disambiguation.
Because of this difference, we leave morphosyntactically congruent ambiguity to future work.

@@

\ex. \label{ex:FSToutput} 
	\a. \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}\\
	    `new'
	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}\\
	    `automaton, sub-machine gun'
%	\c. 
% det/prn (e.g. pos... их, я, мой)
% 'over'generation (e.g. all dets get all gen/nbr/case combinations)
%^^^
%ROB

@@

The constraint grammar is composed of 299 rules which are divided into four
categories: Safe, Safe heuristic, Heuristic, and Syntax labeling. The distribution
of rules is shown in Table~\ref{tab:ruleDist}.

@@

The philosophy is that Safe rules should represent real constraints in
the language. Examples might be that a preposition cannot directly precede a finite verb
or that prepositional case requires a preceding preposition.
